Countering language attrition with PanLex and the Web of Data

### 2014-02-18 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
Minor update

\begin{abstract}
At present, there are approximately 7,000 living languages in the world.
However, some forecasters expect the world to lose most of this linguistic diversity.
The mission of the PanLex project is to investigate whether it is feasible to make all languages intertranslatable and thus usable for world-wide communication, and, if so, whether that new situation will interfere with this expected language attrition.
This mission entails, among other prerequisites, making the lexicons of the world's languages intertranslatable. To accomplish that, the project is systematically integrating and making publicly available all known lexical translations.
Semantic Web technologies can enhance the power of PanLex by flexibly representing and reasoning with the content of its database and interlinking it with linguistic and other resources and annotations.
%Using Semantic Web technologies can support achieving this goal, especially due
% to inference capabilities and the interlinking of the PanLex data with other data sources.
Conversely, PanLex, with its unique collection of translation links between more than a billion pairs of lexemes from more than 9,000 language varieties, can improve the coverage of the Linguistic Web of Data.
In this dataset description paper we detail how we transformed the content of the PanLex database to RDF,
established conformance with the lemon and GOLD data models,
interlinked it with Lexvo and DBpedia, and published it as Linked Data and via
SPARQL.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
At present, there are about 7,000 living languages in the world\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}. As of 25 January 2014, this list contained 6,975 living individual languages.}.
Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.

@@

>>>>

\begin{abstract}
At present, there are approximately 7,000 living languages in the world.
However, some forecasters expect the world to lose most of this linguistic diversity.
The mission of the PanLex project is to investigate whether it is feasible to make all languages intertranslatable and thus usable for world-wide communication, and, if so, whether that new situation will interfere with this expected language attrition.
This mission entails, among other prerequisites, making the lexicons of the world's languages intertranslatable. To accomplish that, the project is systematically integrating and making publicly available all known lexical translations.
Semantic Web technologies can enhance the power of PanLex by flexibly representing and reasoning with the content of its database and interlinking it with
%linguistic and
other resources and annotations.
%Using Semantic Web technologies can support achieving this goal, especially due
% to inference capabilities and the interlinking of the PanLex data with other data sources.
Conversely, PanLex, with its unique collection of translation links between more than a billion pairs of lexemes from more than 9,000 language varieties, can improve the coverage of the Linguistic Web of Data.
In this dataset description paper we detail how we transformed the content of the PanLex database to RDF,
established conformance with the lemon and GOLD data models,
interlinked it with Lexvo and DBpedia, and published it as Linked Data and via
SPARQL.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
At present, there are about 7,000 living languages in the world\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}.
%\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}. As of 25 January 2014, this list contained 6,975 living individual languages.}
%Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, %for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.
Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.

@@

\todo{The combination of user + source document, which has approved any assertion, is the assertion’s source.}
